Summary of Plasma Diagnostic Methods

### 2021-03-12 paper.tex
Updates from Overleaf

% Outline and section completion
% - Magnetic probes
%   - [x] Magnetic field probes
%   - [x] Rogowski coils
% - Electric probes
%   - [x] Langmuir Probes
%   - [ ] Gridded energy analyzers
% - Line radiation spectroscopy
%   - [x] ChERS
%   - [x] LIF/TALIF
%   - [ ] [ ] [ ] Zeeman spectroscopy
% - Scattering of EM radiation
%   - [ ] Coherent Thomson scattering
%   - [x] Incoherent Thomson scattering
% - Imaging
%   - [x] Coded aperture imaging
%   - [x] IR imaging / two-color pyrometry
% - Polarimetry
%   - [ ] Pulsed Polarimetry
%   - [ ] VISAR
% - Fusion products
%   - [x] Neutron Diagnostics
% - Ex Situ Diagnostics
%   - [x] XPS

% Done: 10/17
% Fri
%  - [x] magnetic field probes
%  - [x] rogowski coils
% Sat
%  - [x] neutron diagnostics
%  - [x] ChERS
% Sun
%  - [x] lif/talif
%  - [X] XPS
% Mon
%  - [x] IR/two-color pyrometry
%  - [x] coded aperture imaging
% Tue
%  - [x] incoherent thomson
%  - [x] langmuir probes
% Wed
%  - [ ] Zeeman spectroscopy (THE BIG ONE)
% Thu
%  - [ ] coherent thomson
%  - [ ] visar
% Fri
%  - [ ] gridded energy analyzer
%  - [ ] pulsed polarimetry
%  - [ ] review

@@

Pellet injection is also used in high-energy experiments to introduce localized impurity species. Pellet injectors are frequently used to refuel the burning plasma or to trigger stabilizing edge localized modes \citep{BaylorL.R2015Emwp}. Small pellets are injected at high speed into the plasma where they vaporize, forming a weakly localized ablation cloud. Line radiation from the ablation cloud can then be analyzed for Zeeman splitting measurements. This technique is used in the Tokamak Fusion Test Reactor \citep{doi:10.1063/1.1141775} using lithium pellet injection to determine the magnetic field direction.

@@

Thomson scattering is the elastic scattering of electromagnetic radiation from free charged particles, both electrons and ions.

@@

An incoherent Thomson scattering setup involves a pulsed laser source, an optical collection system with an observation line intersecting the laser path, and a spectrometer capable of resolving narrow wavelength separations. The wavelength of scattered radiation will be Doppler shifted by the thermal motion of the scattering electron, and the intensity of the scattered radiation is proportional to the number of scattering electrons in the scattering volume. Assuming a Maxwellian thermal distribution of electrons, fitting a Gaussian to the scattered radiation spectrum gives the electron temperature. If the spectrometer is capable of sufficiently detailed resolution in frequency space, the spectrum can also be analyzed for non-Maxwellian populations. Recovering the electron density from the scattered intensity reqires an absolute calibration of the diagnostic using a neutral gas.

@@

@@

{\Large TODO \par}
\citep{doi:10.1063/1.1660986}

@@

A promising new magnetic field diagnostic is a LiDAR-like method known as fast-pulsed polarimetry. This method has the potential to resolve some of the most severe limitations of Faraday rotation measurements, providing spatially resolved measurements without the need for inversion of chord-integrated data. The technique combines standard Faraday rotation polarimetry \citep{Pisarczyk1990} and  with a pulsed ranging detection scheme in order to

LiDAR (Light Detection and Ranging) has long been used to generate spatial profiles of distant objects by recording reflections from a pulsed laser source. Devices capable of high power chirped pulse amplification \citep{STRICKLAND1985219} are now available, opening the door to apply pulsed ranging schemes to dense energetic plasmas with millimeter spatial resolution.

@@

\section{Ex Situ Diagnostics}

\subsection{X-ray Photoelectron Spectroscopy (XPS)}

>>>>

% Outline and section completion
% - Magnetic probes
%   - [x] Magnetic field probes
%   - [x] Rogowski coils
% - Electric probes
%   - [x] Langmuir Probes
%   - [ ] Gridded energy analyzers
% - Line radiation spectroscopy
%   - [x] ChERS
%   - [x] LIF/TALIF
%   - [x] [x] [x] Zeeman spectroscopy
% - Scattering of EM radiation
%   - [ ] Coherent Thomson scattering
%   - [x] Incoherent Thomson scattering
% - Imaging
%   - [x] Coded aperture imaging
%   - [x] IR imaging / two-color pyrometry
% - Polarimetry
%   - [ ] Pulsed Polarimetry
%   - [ ] VISAR
% - Fusion products
%   - [x] Neutron Diagnostics
% - Ex Situ Diagnostics
%   - [x] XPS

% Done: 13/18
% Fri
%  - [x] magnetic field probes
%  - [x] rogowski coils
% Sat
%  - [x] neutron diagnostics
%  - [x] ChERS
% Sun
%  - [x] lif/talif
%  - [X] XPS
% Mon
%  - [x] IR/two-color pyrometry
%  - [x] coded aperture imaging
% Tue
%  - [x] incoherent thomson
%  - [x] langmuir probes
% Wed
%  - [x] Zeeman spectroscopy (THE BIG ONE)
% Thu
%  - [ ] coherent thomson
%  - [ ] visar
% Fri
%  - [ ] gridded energy analyzer
%  - [ ] pulsed polarimetry
%  - [ ] review

@@

Pellet injection is also used in high-energy experiments to introduce localized impurity species. Pellet injectors are frequently used to refuel the burning plasma or to trigger stabilizing edge localized modes \citep{BaylorL.R2015Emwp}. Small pellets are injected at high speed into the plasma where they vaporize, forming a roughly localized ablation cloud. Line radiation from the ablation cloud can then be analyzed for Zeeman splitting measurements. This technique is used in the Tokamak Fusion Test Reactor \citep{doi:10.1063/1.1141775} using lithium pellet injection to determine the magnetic field direction.

@@

Thomson scattering is the elastic scattering of electromagnetic radiation from free charged particles in the classical limit. Although the Thomson scattering cross section is extremely small, laser sources of sufficient power ($>100 MW$) to observe Thomson scattering above the background radiation have become commercially available and Thomson scattering measurements have become commonplace in both magnetic fusion and inertial confinement fusion research \citep{doi:10.1063/1.2336451}. Two distinct categories of Thomson scattering measurements can be identified by the product of the wavevector of the scattered radiation and the Debye wavelength. For $k \lambda_{De} \gg 1$, the electrons scatter independently and the scattering is incoherent. When the wavelength is larger than the Debye length, resonant plasma fluctuations form collective effects and the scattering is coherent.

@@

An incoherent Thomson scattering setup involves a pulsed laser source, an optical collection system with an observation line intersecting the laser path, and a spectrometer capable of resolving narrow wavelength separations. The wavelength of scattered radiation will be Doppler shifted by the thermal motion of the scattering electron, and the intensity of the scattered radiation is proportional to the number of scattering electrons in the scattering volume. Assuming a Maxwellian thermal distribution of electrons, fitting a Gaussian to the scattered radiation spectrum gives the electron temperature. If the spectrometer is capable of sufficiently detailed resolution in frequency space, the spectrum can also be analyzed for non-Maxwellian populations. Recovering the electron density from the scattered intensity requires an absolute calibration of the diagnostic using a neutral gas.

@@

Coherent Thomson scattering occurs when the scattering wavelength covers multiple Debye lengths and resonant plasma fluctuations are observed. In this regime, the collective plasma effects can be observed to discern ion  temperature and density information as well as electron temperature. Rather than considering the scattering from an individual electron, the coherent effect must be computed as an integral over many electrons in a finite volume. The full calculation of the scattered power spectrum for a monoenergetic source is given in \citep{https://doi.org/10.1029/JZ065i009p02635}, and can be used to fit the observed scattered spectrum to accurately determine the electron temperature and density. The Thomson-scattered power is proportional to $1 - \sin ^2 \theta \cos ^2 \beta$, where $\theta$ is the angle between the laser and collection optics and $\beta$ is the angle of incident polarization relative to the scattering plane. Observation at an angle of $90^\circ$ with respect to the laser path gives the highest signal intensity, as does polarization of the laser within the scattering plane. Because $k \lambda_{De}$ decreases with decreased angle of incidence, forward scattering may be preferred to reach the coherent regime in low density plasmas where a sufficiently long laser wavelength is not feasible.

While we only observe scattering Thomson scattering from the electrons, because the plasma fluctuations occur over many particles the oscillations will also contain elusive information about the properties of the ions. As long as $Z T_e / T_i > k^2 \lambda _{De} ^2$, collective ion-acoustic features can be observed in the form of two  counter-propagating ion-acoustic waves. The wavelength separation between the ion-acoustic features is approximately
\begin{equation*}
\frac{\Delta \lambda}{\lambda_0} = \frac{4}{c} \sin \left( \frac{\theta}{2} \right) \sqrt{ \frac{ k_b T_e}{m_i} \left[ \frac{Z}{(1 + k ^2 \lambda_{De} ^2)} + \frac{3 T_i}{T_e}\right]}
\end{equation*}
where $m_i$ is the ion mass, $T_i$ is the ion temperature, $T_e$ is the electron temperature, and $Z$ is the ion charge. This is one of the few available direct measurements of the ion temperature. The ion bulk velocity can also be determined by measuring the Doppler shift in these features. In practice, impurities present in the plasma will tend to distort the observed spectrum, making accurate, unambiguous determination of the ion temperature challenging. Even so, the technique is currently in use to measure the ion temperature and velocity in tokamak experiments (ASDEX) and Hall thruster diagnostics (PRAXIS).

@@

A ``Velocity Interferometer System for Any Reflector'' (VISAR) is a diagnostic tool for measuring the velocity of fast-moving surfaces, often used to measure shock propagation in inertial confinement fusion research. As a velocity interferometry system, it computes the phase shift of light reflected from a fast-moving surface, which in turn is used to determine the velocity of the surface. Unlike most velocity interferometry measurements, VISAR is effective for diffuse, un-coated surfaces. The conventional VISAR system (shown in Figure \ref{fig:visar}) is comprised of a pair of wide-angle Michelson interferometers operating on different polarizations of light reflected from the moving surface. A wave plate in one leg of the interferometer adds a $90^\circ$ phase shift between the two polarizations. A third detector monitors the raw intensity of the input light.
\begin{figure}
  \centering
  \includegraphics[width=0.6\textwidth]{visar.pdf}
  \caption{Diagram of the conventional VISAR setup, showing beam intensity monitor (BIM), beam splitter (BS), 1/8 wave plate (WP), etalon (E), mirrors (M1, M2), polarizing beam splitter (PBS), and detectors (D1, D2). From \citep{10.2172/886901}.}
\label{fig:visar}
\end{figure}
In this configuration, the phase of the input light $\phi(t)$ can be computed by plotting the intensities measured at the detectors $D1$ and $D2$ normalized by the beam intensity monitor signal on an ellipse defined by:
\begin{equation*}
D_x \equiv \frac{D_1(t)}{D_{BIM} (t)} \qquad D_y \equiv \frac{D_2 (t)}{D_{BIM}(t)} \qquad \tan \phi (t) = \tan \epsilon + \frac{D_y (t) - y_0 }{D_x (t) - x_0} A \sec \epsilon
\end{equation*}
where $x_0$, $y_0$, $A$, and $\epsilon$ are independent of the input power. Because the interferometer signal is normalized by the beam intensity monitor, the setup is insensitive to variations of the input intensity. This is what allows for velocity measurements of diffuse surfaces. The expression for $\tan \phi$ (in contrast to $\cos \phi$ for a single Michelson interferometer) makes it possible to remove the sign ambiguity from $\phi$ and distinguish velocities towards and away from the detector.

One drawback to the conventional VISAR setup is that only half of the incident light reaches the detectors, while the rest is lost to the opposite side of the beam splitter. We can add an additional VISAR on the other side of the beam splitter in place of the beam intensity monitor, generating a second set of signals very similar to the first pair. By pairwise subtracting the signals using electronic differential amplifiers (to impart a $180^\circ$ phase shift), the result is known as a push-pull VISAR. The phase of the input light can then be calculated from the resulting pair of signals in the same manner as a conventional VISAR. The additional light efficiency can improve the noise efficiency of the VISAR by approximately $40\%$ over a conventional VISAR, and if properly balanced is insensitive to variations in the input intensity and incoherent light \citep{10.2172/886901}. The push-pull VISAR has largely supplanted conventional VISAR in most applications.

@@

A promising new magnetic field diagnostic is a LIDAR-like method known as fast-pulsed polarimetry. This method has the potential to resolve some of the most severe limitations of Faraday rotation measurements, providing spatially resolved measurements without the need for inversion of chord-integrated data. The technique combines standard Faraday rotation polarimetry \citep{Pisarczyk1990} and  with a pulsed ranging detection scheme in order to

LIDAR (Light Detection and Ranging) has long been used to generate spatial profiles of distant objects by recording reflections from a pulsed laser source. Devices capable of high power chirped pulse amplification \citep{STRICKLAND1985219} are now available, opening the door to apply pulsed ranging schemes to dense energetic plasmas with millimeter spatial resolution.

@@

\section{X-ray Photoelectron Spectroscopy (XPS)}
